---
title: "HW 4 - College Majors"
author: "Annie Hou"
date: "02/19/2024"
output: html_document
---

## Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse)
library(scales)
library(fivethirtyeight)
```

## Exercises

### Exercise 1

The top-3 majors that have the highest percentage of women are: Early Childhood Education, Communication Disorders Sciences And Services, and Medical Assisting Services.

```{r ex1}
college_recent_grads %>%
  arrange(desc(sharewomen)) %>%
  select(major, total, sharewomen) %>%
  top_n(3, sharewomen)

```

### Exercise 2

We choose the median because it is less likely to be affected by outliers. For example, a few people may have extremely high incomes, which would increase the mean a lot; or a few people may also have extremely low incomes, leading to a large decrease in mean. However, the median only cares about the middle value of a series of data, which will not be largely affected by those extremes. Therefore, median could better indicate the typical income for a group.


### Exercise 3

I chose a binwidth of $1000, because it offers a more detailed view of the means. It could show more specific patterns in the data, which could be helpful for more in-depth analysis.

```{r median}
ggplot(college_recent_grads, aes(x = median)) +
  geom_histogram(binwidth = 1000)
```

```{r summary-stats}
college_recent_grads %>%
  summarize(min = min(median), max = max(median),
            mean = mean(median), med = median(median),
            sd = sd(median), 
            q1 = quantile(median, probs = 0.25),
            q3 = quantile(median, probs = 0.75))
```

### Exercise 4

The center of the distribution is 36000, which is lower than the mean of 40151, which means the distribution of the incomes are clustered on the left side. It also means that while the majority majors have the median incomes that are clustered on the lower end, there are still some majors with significantly higher median incomes, like the plot shows, and that makes the mean larger than the median.

The spread of the data is 11470, measured by the standard deviation. It means the distribution of median incomes is relatively wide, which we can also oberserve from the graph. The interquartile range is from the first quartile at 33000 to the third quartile at 45000, which includes the middle 50% of the data, so half of the majors have median incomes within this range.

For the unusal observations, we can see from the graph that there are majors that have very high median incomes far on the right side. By the max value in the table, we know that it has reached 110000, which is far more than the median and mean. Such outlier could affect the mean and spread of the distribution.

### Exercise 5

```{r facet-major}
ggplot(college_recent_grads, aes(x = median)) +
  geom_histogram(binwidth = 1000) +
  facet_wrap(~major_category) + 
  theme_minimal() +
  labs(title = "Distribution of Median Income by Major Category",
       x = "Median Income",
       y = "Count")
```


### Exercise 6

```{r highest}
college_recent_grads %>%
  group_by(major_category) %>%
  summarize(median_income = median(median)) %>%
  arrange(desc(median_income))
```

From the table, we can see that the top 3 majors with the highest typical median income are: engineering, computers & mathematics, and business.

### Exercise 7

```{r least}
college_recent_grads %>%
  count(major_category) %>%
  arrange(n)
```

From the table, we can see that the least 3 popular majors are: interdisciplinary, communications & journalism, law & public policy.

### Exercise 8

```{r stem}
stem_categories <- c("Biology & Life Science",
                     "Computers & Mathematics",
                     "Engineering",
                     "Physical Sciences")
college_recent_grads <- college_recent_grads %>%
  mutate(major_type = ifelse(major_category %in% stem_categories,
                             "stem", "not stem"))
college_recent_grads %>%
  filter(major_type == "stem", median < 36000)%>%
  select(major, median, p25th, p75th) %>%
  arrange(desc(median))

```


### Exercise 9

```{r plot-women}
ggplot(college_recent_grads, aes(x = sharewomen, y = median, color = major_type)) +
  geom_point() +
  labs(x = "Proportion of Women", y = "Median Income", color = "Major Type") +
  theme_minimal() +
  ggtitle("Scatterplot of Median Income vs. Proportion of Women, Colored by STEM Field")

```

The scatterplot first shows a negative correlation between the proportion of women and the median income of that major. Because when the proportion of women increases (to the right side of the x-axis), the median income tends to decrease. Also, there are less STEM majors as the proportion of women increases and the median income decreases, because non-STEM majors (red dots) are more clustered at the lower median income range and have a higher proportion of women (right bottom side of the plot). This may indicate that majors with more women may less likely to have high median income or being STEM. 

### Exercise 10

```{r unemployment}
ggplot(college_recent_grads, aes(x = sharewomen, y = unemployment_rate)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Scatterplot of Unemployment Rate vs. Proportion of Women in Major",
       x = "Proportion of Women",
       y = "Unemployment Rate") +
  theme_minimal()
```

I choose to investigate the relationship between the proportion of women in a major and the unemployment rate of that major. The resulting scatterplot shows a non-linear relationship that tends to go up for the first half and go down for the second half. As the proportion of women increases, the unemployment rate initially appears to decrease slightly and then increases after a certain point. It means that the unemployment rate is correlated to the proportion of women, but the direction changes after certain threshold. The shaded area around the line indicates the confidence interval, which widens as the beginning and the ending part of the curve, which implies more variability in the unemployment rate among majors with a very low or high proportion of women. Therefore, the proportion of women is correlated to the unemployment rate.